On the Galaxy Size--Halo Connection

### 2017-08-30 paper/galsize_paper.tex
Finished sketch of model

%%%%%%%%  galaxy properties  %%%%%%%%
\newcommand{\rhalf}{R_{1/2}}
\newcommand{\rhalfdisk}{R_{1/2}^{\rm disk}}
\newcommand{\rhalfbulge}{R_{1/2}^{\rm bulge}}
\newcommand{\adisk}{A_{\rm disk}}
\newcommand{\abulge}{A_{\rm bulge}}
\newcommand{\alphadisk}{\alpha_{\rm disk}}
\newcommand{\alphabulge}{\alpha_{\rm bulge}}
\newcommand{\rvir}{R_{\rm vir}}
\newcommand{\bt}{{\rm B/T}}
\newcommand{\mstar}{M_{\ast}}
\newcommand{\ssfr}{{\rm sSFR}}
\newcommand{\sfr}{{\rm SFR}}

@@

\begin{abstract}
We derive empirical modeling constraints on the connection between dark matter halos and the half-mass radius $\rhalf$ of galaxy bulges and disks. We show that both $\rhalfdisk$ and $\rhalfbulge$ are well-described by power law scaling relations with halo virial radius, $\rhalf=A\rvir^{\alpha}.$ Novel to this work, we use new SDSS measurements of the $\rhalf-$dependence of galaxy clustering to constrain the model parameters, $A_{\rm bulge}, A_{\rm disk}, \alpha_{\rm bulge},\alpha_{\rm disk},$ and log-normal scatter $\sigma_{R_{1/2}}.$ Even when only coarsely tuning these parameters to the observed one-point functions $\langle\rhalfdisk|\mstar^{\rm disk}\rangle$ and $\langle\rhalfbulge|\mstar^{\rm bulge}\rangle,$ our model accurately predicts the observed two-point clustering on small- and large-scales. This success non-trivial, as we show that galaxy clustering is highly sensitive to the physics that shapes satellite galaxy profiles. We find no evidence for the commonly assumed relation between halo spin $\halospin$ and $\rhalfdisk,$ and show that this assumption cannot be meaningfully constrained with either the clustering or lensing of  $L_{\ast}$ galaxies. Our results provide simple boundary conditions for more complex and fine-grained models of galaxy size. We make our python code publicly available to support cosmological surveys that require realistic synthetic galaxy populations.
\end{abstract}

@@

We additionally use the catalog of SDSS galaxy profile decompositions provided by \citet{meert_etal15}. In the version of this catalog that we use, two-dimensional $r-$band profiles were fit with a two-component de Vaucouleurs + exponential profile for the bulge and disk, respectively. We cross-match our DR10 sample against the \citet{meert_etal15} scale radii and flux ratios of the two components, giving us a volume-limited galaxy sample with values of $\mstar,\ssfr,\rhalfbulge,\rhalfdisk,$ and $\bt.$

@@

>>>>

%%%%%%%%  galaxy properties  %%%%%%%%
\newcommand{\rhalf}{R_{1/2}}
\newcommand{\rhalfdisk}{R_{1/2}^{\rm disk}}
\newcommand{\rhalfbulge}{R_{1/2}^{\rm bulge}}
\newcommand{\adisk}{A_{\rm disk}}
\newcommand{\abulge}{A_{\rm bulge}}
\newcommand{\alphadisk}{\alpha_{\rm disk}}
\newcommand{\alphabulge}{\alpha_{\rm bulge}}
\newcommand{\sigmarhalf}{\sigma_{\rm R_{1/2}}}
\newcommand{\rvir}{R_{\rm vir}}
\newcommand{\bt}{{\rm B/T}}
\newcommand{\mstar}{M_{\ast}}
\newcommand{\ssfr}{{\rm sSFR}}
\newcommand{\sfr}{{\rm SFR}}

@@

\begin{abstract}
We derive empirical modeling constraints on the connection between dark matter halos and the half-mass radius $\rhalf$ of galaxy bulges and disks. We show that both $\rhalfdisk$ and $\rhalfbulge$ are well-described by power law scaling relations with halo virial radius, $\rhalf=A\rvir^{\alpha}.$ Novel to this work, we use new SDSS measurements of the $\rhalf-$dependence of galaxy clustering to constrain the model parameters, $A_{\rm bulge}, A_{\rm disk}, \alpha_{\rm bulge},\alpha_{\rm disk},$ and log-normal scatter $\sigmarhalf.$ Even when only coarsely tuning these parameters to the observed one-point functions $\langle\rhalfdisk|\mstar^{\rm disk}\rangle$ and $\langle\rhalfbulge|\mstar^{\rm bulge}\rangle,$ our model accurately predicts the observed two-point clustering on small- and large-scales. This success non-trivial, as we show that galaxy clustering is highly sensitive to the physics that shapes satellite galaxy profiles. We find no evidence for the commonly assumed relation between halo spin $\halospin$ and $\rhalfdisk,$ and show that this assumption cannot be meaningfully constrained with either the clustering or lensing of  $L_{\ast}$ galaxies. Our results provide simple boundary conditions for more complex and fine-grained models of galaxy size. We make our python code publicly available to support cosmological surveys that require realistic synthetic galaxy populations.
\end{abstract}

@@

We additionally use the catalog of SDSS galaxy profile decompositions provided by \citet{meert_etal15}. In the version of this catalog that we use, two-dimensional $r-$band profiles were fit with a two-component de Vaucouleurs + exponential profile for the bulge and disk, respectively. We cross-match our DR10 sample against the \citet{meert_etal15} scale radius of the composite profile and flux ratios of the two components, giving us a volume-limited galaxy sample with values of $\mstar,\ssfr,\rhalf,$ and $\bt.$

@@

Motivated by the \citet{kravtsov13} results, we model the half-mass radii of galaxies as separate power law functions of halo virial radius:

\begin{eqnarray}
\rhalfbulge &=& \abulge\rvir^{\alphabulge} \\
\rhalfdisk &=& \adisk\rvir^{\alphadisk} 
\end{eqnarray}

For central galaxies residing in host halos, we use the present-day virial radius; for satellite galaxies we use the virial radius at the time of accretion, using physical units of $\kpc.$ We model the composite size of model galaxies as 
\beq
\rhalf = (\bt)\rhalfbulge + (1-\bt)\rhalfdisk.
\eeq

When generating Monte Carlo realizations of our model galaxy population, we add log-normal scatter $\sigmarhalf$ separately to each component. Our model for sizes thus has five free parameters: $\abulge, \adisk, \alphabulge, \alphadisk,$ and $\sigmarhalf.$
